Fast Sorting on Modern Computers

### 2005-05-27 paper/docs/conclusions.tex
make way for the new and improved fs
removed files not for paper removed extra bits
done some spellchecking

The quicksort results were also very similar, though the multi-quicksort results
had a slightly different shape to LaMarca's. However, the instruction count
results for each quicksorts agreed with LaMarca's results, as did cache miss
results of the base and cache-conscious versions.

@@

\section{Best Sort}
The results also helped answer the question of which is the best sort. However,
it is important to remember that sorts have different purposes and properties,
so this shall be considered. A comparative chart of the fastest sorts is shown
in Figure \ref{All cycles}. This shows the cycles per key of each of the major
sorts and their derivatives, measured using Pentium 4 performance counters.
Neither the elementary sorts nor heapsorts are on the chart, as their results
are significantly worse than the other results, which makes it difficult to
accurately see those results.

@@

The best stable sort is radixsort. The best comparison-based stable sort is
double tiled mergesort. The best in-place sort is quicksort. The best in-place
stable sort is, sadly, insertion sort. The best elementary sort is also
insertion sort; as such it should be only used for very small data sets.  The
best sort for very large data sets is radixsort. The best overall sort was
quicksort, though radixsort is very close, and the radixsort considered here was
not optimised. It is expected that an optimised radixsort would beat quicksort
for smaller sets.

>>>>

The quicksort results were also very similar, though the multi-quicksort results
had a very different shape to LaMarca's. However, the instruction count
results for each quicksort agreed with LaMarca's results, as did cache miss
results of the base and cache-conscious versions.

@@

\section{Best Sort}
The results also helped answer the question of which is the best sort. However,
it is important to remember that sorts have different purposes and properties,
so this shall be considered. A comparative chart of the fastest sorts is shown
in Figure \ref{All cycles}. This shows the cycles per key of each of the major
sorts and their derivatives, measured using Pentium 4 performance counters.
None of the elementary sorts or heapsorts are on the chart, as their results
are significantly worse than the other results, which makes it difficult to
accurately see them.

@@

The best stable sort is radixsort. The best comparison-based stable sort is
double tiled mergesort. The best in-place sort is quicksort. The best in-place
stable sort is, sadly, insertion sort. The best elementary sort is also
insertion sort; as such it should be only used for very small data sets.  The
best sort for very large data sets is radixsort. The best overall sort was
quicksort, though radixsort is very close, and the radixsort considered here was
not optimized. It is expected that an optimized radixsort would beat quicksort
for smaller sets.
% TODO redo above paragraph

### 2005-05-27 paper/docs/heap.tex
done some spellchecking

A difficulty of heapsort is the need to do a bounds check every time. In a heap,
every parent must have either no children, or two children. The only exception
to this is the last parent. It is possible that this has only one child, and it
is necessary to check for this as a result. This occurs every two heapifies, as
one key is removed from the heap every step. Therefore it is not possible to just
pretend the heap is slightly smaller, and sift up the final key later, as it
might be with a staticly sized heap.

@@

In the graph, the base heapsort results flat-line for a step longer than the
other heapsorts' results. This is due to being in-place, since twice as many
keys fit in the cache without a corresponding increase in misses. For the same
reason, the base heapsorts have half as many level 2 cache misses of the later,
out-of-place sorts (See Figure \ref{heap level 2 misses}).

>>>>

A difficulty of heapsort is the need to do a bounds check every time. In a heap,
every parent must have either no children, or two children. The only exception
to this is the last parent. It is possible that this has only one child, and it
is necessary to check for this as a result. This occurs every two heapifies, as
one key is removed from the heap every step. Therefore it is not possible to just
pretend the heap is slightly smaller, and sift up the final key later, as it
might be with a statically sized heap.

@@

In the graph, the base heapsort's results flat-line for a step longer than the
other heapsorts' results. This is due to being in-place, since twice as many
keys fit in the cache without a corresponding increase in misses. For the same
reason, the base heapsorts have half as many level 2 cache misses of the later,
out-of-place sorts (See Figure \ref{heap level 2 misses}).

### 2005-05-27 paper/docs/merge.tex
done some spellchecking

The final impriovement was to fully sort keys in the level 1 cache, before
continuing to level 2 cache as in the tiled mergesort. The expected improvement
is not as significant as in the case of the level 2 cache, since is the penalty
for level 1 misses is not great. For this reason, this sort is called Double
Tiled Mergesort instead of Tiled Mergesort, however, it is referred to as both
in this report.

@@

The reuslts show that the seconsary loops are always taken. Taken here means broken out of, hence seldom used. The main action occurs in N3, the main comparator, which has a 50\% prediction ratio.

@@

Figure c shows base mergesort. This shows a lower number of branches and a higher precentage of correct predictions. The lower number of branches is probably due to the presort. there is a higher percentage of correct predictions, but the actual number of correct predictions in the same. This implies that doing the presort removes incorrect predictions (or, more to the point, moves incorrect predictions to another part of the algorithm). Note too, that both sides are used equally, which doesnt happen later. Two thirds of the branches are not-taken, which means loops execute twice before breaking. this is the same results that we found in quicksort later. Im sure knuth explains this somewhere in the fine print.

>>>>

The final improvement was to fully sort keys in the level 1 cache, before
continuing to level 2 cache as in the tiled mergesort. The expected improvement
is not as significant as in the case of the level 2 cache, since is the penalty
for level 1 misses is not great. For this reason, this sort is called Double
Tiled Mergesort instead of Tiled Mergesort, however, it is referred to as both
in this report.

@@

The results show that the secondary loops are always taken. Taken here means broken out of, hence seldom used. The main action occurs in N3, the main comparator, which has a 50\% prediction ratio.

@@

Figure c shows base mergesort. This shows a lower number of branches and a higher percentage of correct predictions. The lower number of branches is probably due to the presort. there is a higher percentage of correct predictions, but the actual number of correct predictions in the same. This implies that doing the presort removes incorrect predictions (or, more to the point, moves incorrect predictions to another part of the algorithm). Note too, that both sides are used equally, which doesn't happen later. Two thirds of the branches are not-taken, which means loops execute twice before breaking. this is the same results that we found in quicksort later. I'm sure knuth explains this somewhere in the fine print.
